Compositional equivalences based on Open pNets

### 2020-02-05 2016-Weak-FH-Bisim/WeakBisim.tex
remove *pdf from the repo
full WOT in page 22

    \begin{figure}[ht]
  %\centerline{\includegraphics[width=10cm]{ATG/SPImplOpen}}
  \caption{Open Automaton of the Simple Protocol Implementation}  \label{SimpleProtCounter:ImplOA}
\end{figure}

@@

 WOT2 here is the transition delivering the result of the algorithm to hole $Q$. Remark that it includes a loop increasing the error counter, occurring before the action of $Q$, and this action is using the resulting value of the counter. Let us go into the details, using rules WT1/2/3.  First WT1 adds a $\tau$ loop on each state. Then consider application of Rule WT3 on a sequence 3 internal actions ``$\tau\subst{ec\gets ec+1};\tau\subst{ec\gets ec+1};\tau$''. Hole actions and predicates are empty, we just concentrate on the Post operations. Rule WT2 transforms each 3 OTs into WOTs. In the drawing we use \texttt{ec++} to abbreviate the Post: $\subst{ec\gets ec+1}$. Then WT3 produces the weak OT:
  $\openrule
         {\{\}, True,
			\subst{\texttt{ec++}\shortotimes\texttt{ec++}\shortotimes\emptyset} }
         {s \OTWeakarrow {\tau\subst{\texttt{ec++}}} s'}$
         that you can rewrite as:      
  $\openrule
         {\{\}, True,
			\subst{ec\gets ec+2} }
         {s \OTWeakarrow \tau s'}$

@@

         For any of these, apply again WT3, using the central \texttt{q-recv(ec)} transition as $\alpha$, and you get:
      $\openrule
         {\{Q\mapsto\texttt{q-revc(ec)}\subst{ec\gets ec+n}\}, True,
			\subst{ec\gets ec+n} }
         {s \OTWeakarrow {\texttt{out(m,ec)}\subst{ec\gets ec+n}} s'}$
\end{example}

@@

\begin{figure}[h]
   \centerline{\includegraphics[width=15cm]{ATG/SPSpecWeakOpen}}
  \caption{Weak Open Automaton of the Specification \TODO{out of date}}
   \label{SimpleProtCounter:WeakSpecOA}
\end{figure}

\begin{figure}[h]
   \centerline{\includegraphics[width=10cm]{XFIG/SPSpecWeakOpen}}
  \caption{Weak Open Automaton of the Specification}
   \label{SimpleProtCounter:WeakSpecOA}
\end{figure}

@@

\TODO{Eric: insert here a table with the details of the WOTs.Il manque encore au moins 3 WOT dans ce dessin...}

\smallskip
$1^st$ try, user-friendly style:

$$ 3 = \tau^*; \underline{in(m)}\{s1\_c:=0, s1\_m:=m\}; \tau\{s1\_c++\}^*$$

\smallskip
$2^nd$ try, full formal style ?:
\medskip

Here we denote as $post_n$ the effect of the Strong transition $OT_n$ from Figure \ref{}:
\smallskip

@@

$post_3 = (s1\_c\gets 0, s1\_m\gets m)$

$post_4 = (m1\_msg\gets s1\_msg, m1\_c\gets s1\_c)$

$post_5 = (m2\_c\gets m1\_c)$

$post_6 = (s1\_c\gets m2\_c+1)$

$post_7 = (r1\_msg\gets m1\_msg, r1\_c\gets m1\_c)$

$post_8 = ()$
\medskip

Then the effect of one single $100 -> 210 -> 220 -> 100$ loop as:
\smallskip

$Post_{456} = post_6\shortotimes post_5\shortotimes post_4$
\medskip

And we can now define the weak OTs:
\smallskip

$ 1 = \openrule
{\{P\mapsto p\_a\}, [\forall m. p\_a \neq p\_send(m)], ()}
{000 \OTWeakarrow {\tau} 000}
$

$ 3 = \openrule
         {\{P\mapsto p\_send(m)\}, True, post_4\shortotimes {Post_{456}}^*\shortotimes\ post_3}
         {000 \OTWeakarrow {\texttt{in(m)}} 100}$

$ 4 = \openrule
         {\{\}, True, post_4\shortotimes {Post_{456}}^* }
         {100 \OTWeakarrow {\tau} 210}
         $

$ 5 = \openrule
         {\{\}, True, {Post_{564}}^* }
         {210 \OTWeakarrow {\tau} 220}
         $

$ 6 = \openrule
         {\{\}, True, {Post_{456}}^*\shortotimes post_6 }
         {220 \OTWeakarrow {\tau} 100}
         $

$ 7 = \openrule
         {\{\}, True, post_7\shortotimes {Post_{564}}^* }
         {220 \OTWeakarrow {\tau} 100}
         $

@@

@@

The first piece of code is the textual definition of the SimpleProtSec
pNet, that was drwan in Figure \ref{SimpleProt:Impl}, page \pageref{SimpleProt:Impl}. This code should be intuitive enough to read, with the following
conventions, that brings some user-friendly features that are maps by the editor into pure pNet constructs.

>>>>

    \begin{figure}[ht]
  %\centerline{\includegraphics[width=10cm]{ATG/SPImplOpen}}
  \caption{Open Automaton of the Simple Protocol Implementation}  %\label{SimpleProtCounter:ImplOA}
\end{figure}

@@

 WOT2 here is the transition delivering the result of the algorithm to hole $Q$. Remark that it includes a loop increasing the error counter, occurring before the action of $Q$, and this action is using the resulting value of the counter. Let us go into the details, using rules WT1/2/3.  First WT1 adds a $\tau$ loop on each state. Then consider application of Rule WT3 on a sequence 3 internal actions ``$\tau\subst{ec\gets ec+1};\tau\subst{ec\gets ec+1};\tau$''. Hole actions and predicates are empty, we just concentrate on the Post operations. Rule WT2 transforms each 3 OTs into WOTs.  Then WT3 produces the weak OT:
  $\openrule
         {\{\}, True,
			\subst{\texttt{ec++}\shortotimes\texttt{ec++}\shortotimes\emptyset} }
         {s \OTWeakarrow {\tau\subst{\texttt{ec++}}} s'}$
         that you can rewrite as:      
  $\openrule
         {\{\}, True,
			\subst{ec\gets ec+2} }
         {s \OTWeakarrow \tau s'}$

@@

         For any of these, apply again WT3, using the central \texttt{q\_recv(ec)} transition as $\alpha$, and you get:
      $\openrule
         {\{Q\mapsto\texttt{q\_revc(ec)}\subst{ec\gets ec+n}\}, True,
			\subst{ec\gets ec+n} }
         {s \OTWeakarrow {\texttt{out(m,ec)}\subst{ec\gets ec+n}} s'}$
\end{example}

@@

\begin{figure}[h]
   \centerline{\includegraphics[width=13cm]{ATG/SPSpecWeakOpen}}
  \caption{Weak Open Automaton of the Specification \TODO{out of date}}
   \label{SimpleProtCounter:WeakSpecOA}
\end{figure}

\begin{figure}[h]
   \centerline{\includegraphics[width=12cm]{XFIG/SPSpecWeakOpen}}
  \caption{Weak Open Automaton of the Specification}
   \label{SimpleProtCounter:WeakSpecOA}
\end{figure}

@@

%% \TODO{Eric: insert here a table with the details of the WOTs.Il manque encore au moins 3 WOT dans ce dessin...}

%% \smallskip
%% $1^st$ try, user-friendly style:

%% $$ 3 = \tau^*; \underline{in(m)}\{s1\_c:=0, s1\_m:=m\}; \tau\{s1\_c++\}^*$$

%% \smallskip
%% $2^nd$ try, full formal style ?:
%% \medskip

Now let us give some details about the construction of the weak automation of the implementation pNet, obtained by application of the weak rules as explained above. Here we denote as $post_n$ the effect (= susbtitution) of the Strong transition $OT_n$ from Figure \ref{SimpleProtCounter:ImplOA}:
\smallskip

@@

$post_3 = (\texttt{s1\_c}\gets 0, \texttt{s1\_m}\gets \texttt{m})$

$post_4 = (\texttt{m1\_msg}\gets \texttt{s1\_msg}, \texttt{m1\_c}\gets \texttt{s1\_c})$

$post_5 = (\texttt{m2\_c}\gets \texttt{m1\_c})$

$post_6 = (\texttt{s1\_c}\gets \texttt{m2\_c+1})$

$post_7 = (\texttt{r1\_msg}\gets \texttt{m1\_msg}, \texttt{r1\_c}\gets \texttt{m1\_c})$

$post_8 = post_9 = ()$
\medskip

Then the effect of one single $100 \xrightarrow{OT_4} 210 \xrightarrow{OT_5} 220 \xrightarrow{OT_6} 100$ loop is:
\smallskip

$Post_{456} = post_6 \shortotimes\ post_5 \shortotimes\ post_4
= (s1\_c\gets s1\_c+1)$.
\medskip

So if we denote ${Post_{456}}^*$ any iteration of this loop, we get ${Post_{456}}^* = (\texttt{s1\_c}\gets \texttt{s1\_c}+n)$ for any $n\ge 0$, and we will get e.g. the Post of the weak OT number 3' as $ post_4\shortotimes {Post_{456}}^*\shortotimes\ post_3 = (\texttt{m1\_msg}\gets m, \texttt{m1\_c}\gets n), \forall n\ge 0$
\medskip

We can now show some of the weak OTs of Figure \ref{SimpleProtCounter:SpecOA-partial} (the full table will be included in Appendix \ref{Appendix:FullExample}) :
\smallskip

$ 1 = \openrule
{\{\texttt{P}\mapsto \texttt{p\_a}\}, [\forall \texttt{m}. \texttt{p\_a} \neq \texttt{p\_send(m)} ], ()}
{000 \OTWeakarrow {\tau} 000}$

$ 2 = \openrule
{\{\texttt{Q}\mapsto \texttt{q\_b}\}, [\forall \texttt{m}. \texttt{q\_b} \neq \texttt{q\_recv(m)} ], ()}
{000 \OTWeakarrow {\tau} 000}$

$ 3 = \openrule
  {\{\texttt{P}\mapsto \texttt{p\_send(m)}\}, True,
    (\texttt{s1\_msg}\gets m, \texttt{s1\_c}\gets n)}
  {000 \OTWeakarrow {\underline{\texttt{in(m)}}} 100}
  \forall n\ge 0$

$ 3' = \openrule
  {\{\texttt{P}\mapsto \texttt{p\_send(m)}\}, True,
    (\texttt{m1\_msg}\gets m, \texttt{m1\_c}\gets n)}
  {000 \OTWeakarrow {\underline{\texttt{in(m)}}} 210}
  \forall n\ge 0$

$ 3'' = \openrule
  {\{\texttt{P}\mapsto \texttt{p\_send(m)}\}, True, 
    (\texttt{m2\_c}\gets n)}
  {000 \OTWeakarrow {\underline{\texttt{in(m)}}} 220}
  \forall n\ge 0$

         $ 4 = \openrule
         {\{\}, True, (\texttt{m1\_msg}\gets \texttt{s1\_msg}, \texttt{m1\_c}\gets \texttt{s1\_c}+n)}
         {100 \OTWeakarrow {\tau} 210}
         \forall n\ge 0$

$ 5 = \openrule
         {\{\}, True, (\texttt{m2\_c}\gets \texttt{m1\_c}+n)}
         {210 \OTWeakarrow {\tau} 220}
         \forall n\ge 0$

$ 6 = \openrule
         {\{\}, True, (\texttt{s1\_c}\gets \texttt{m2\_c}+n)}
         {220 \OTWeakarrow {\tau} 100}
         \forall n\ge 0$

$ 7 = \openrule
         {\{\}, True, (\texttt{r1\_msg}\gets \texttt{s1\_msg}, \texttt{r1\_c}\gets \texttt{s1\_c}+n)}
         {100 \OTWeakarrow {\tau} 201}
         \forall n\ge 0$

$ 7' = \openrule
         {\{\}, True, (\texttt{r1\_msg}\gets \texttt{m1\_msg}, \texttt{r1\_c}\gets \texttt{m1\_c}+n)}
         {210 \OTWeakarrow {\tau} 201}
         \forall n\ge 0$

$ 7'' = \openrule
         {\{\}, True, (\texttt{r1\_msg}\gets \texttt{m1\_msg}, \texttt{r1\_c}\gets \texttt{m2\_c}+n)}
         {220 \OTWeakarrow {\tau} 201}
         \forall n\ge 1$

$ 8 = \openrule
         {\{\}, True, ()}
         {201 \OTWeakarrow {\underline{\texttt{out(m,c)}}\subst{(m\gets \texttt{r1\_msg}, c\gets \texttt{r1\_c}}} 202}
         $

@@

\label{Appendix:FullExample}

@@

The first piece of code is the textual definition of the SimpleProtSec
pNet, that was drawn in Figure \ref{SimpleProt:Impl}, page \pageref{SimpleProt:Impl}. This code should be intuitive enough to read, with the following
conventions, that brings some user-friendly features that are maps by the editor into pure pNet constructs.
